Gravitational Lens Modeling in a Citizen Science Context

### 2014-07-27 text/ms.tex
Reordered authors to show properly Prasenjit's role, and to give the modellers full credit. Fixes #20. Also beefed up abstract a bit, addressing #19

\definecolor{purple}{RGB}{150,0,200}
\newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf Phil: #1}}

@@

\title{Lens Modeling in a Citizen-Science context}

\author[Kueng et al]{Rafael Kueng,$^{1}$
Prasenjit Saha$^{1}$
Elisabeth Baeten,$^{5}$
Claude Cornen,$^{5}$
\newauthor
Christine McMillan,$^{5}$
Julianne Wilcox,$^{5}$
Jonas Odermatt,$^{6}$
Jonathan Coles,$^{7}$
\newauthor
Phil Marshall,$^{2}$
Anupreeta More,$^{3}$
Aprajita Verma,$^{4}$
and Surhud More$^{3}$ \\
$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
$^{2}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
$^{3}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
$^{4}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
$^{5}$Spacewarps\\
$^{6}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
$^{7}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France}
%$^{8}$Adler Planetarium, 1300 S. Lake Shore Drive, Chicago, IL 60605, USA\\

@@

\begin{abstract}
We develop a method to enable collaborative modelling of gravitational
lenses and lens candidates by experienced but non-professional lens
enthusiasts.  It uses an existing free-form modelling program (GLASS)
but provides the modelling input in a novel way, via a user-provided
diagram that is essentially a sketch of an arrival-time surface.

An implementation (\spl) is tested in a modelling challenge using 29
simulated lenses out of a larger set created for a citizen-science
lens search (\sw).  We find that amateur lensers inferred the image
parities and time ordering consistently in some lenses, while tending
to make errors in other lenses, depending on the image morphology.
Errors in image parity and time ordering lead to large errors in the
mass distribution, but the enclosed mass is more robust.  Shortcomings
include (a)~lens profiles that tend to be systematically shallower
than the lenses, and consequent over-estimation of the Einstein radius
by $\sim25\%$, and (b)~lack of detailed source modelling.  Ideas for
improvement are discussed.
\end{abstract}

>>>>

\definecolor{purple}{RGB}{150,0,200}
\newcommand{\phil}[1]{\noindent\textcolor{purple}{\bf #1}}

@@

\title{Gravitational Lens Modeling in a Citizen Science Context}

\author[Kueng et al]{Rafael Kueng,$^{1}$
Prasenjit Saha,$^{1}$
Elisabeth Baeten,$^{2}$
Claude Cornen,$^{2}$
\newauthor
Christine McMillan,$^{2}$
Julianne Wilcox,$^{2}$
Jonas Odermatt,$^{3}$
Jonathan Coles,$^{4}$
\newauthor
Aprajita Verma,$^{5}$
Anupreeta More,$^{6}$
Surhud More,$^{6}$
and Phil Marshall$^{7}$ \\
$^{1}$Physik-Institut, University of Zurich, 190 Winterthurerstrasse, 8057, Zurich, Switzerland\\
$^{2}$The Space Warps Collaboration\\
$^{3}$Kantonsschule Zug, L\"ussiweg 24, 6300 Zug, Switzerland\\
$^{4}$Exascale Computing Lab, Versailles Saint-Quentin-en-Yveline University, 55 Avenue des Etats Unis, 78000, Versailles, France\\
$^{5}$Oxford Astrophysics, Keble Road, Oxford OX1 3RH, UK\\
$^{6}$Kavli Institute for the Physics and Mathematics of the Universe, University of Tokyo, 5-1-5 Kashiwanoha, Kashiwa-shi 277-8583, Japan\\
$^{7}$Kavli Institute for Particle Astrophysics and Cosmology, Stanford University, 452 Lomita Mall, Stanford, CA 94035, USA\\
}

@@

\begin{abstract}

We develop a method to enable collaborative modelling of gravitational
lenses, and lens candidates, by non-professional lens enthusiasts.  It
uses an existing free-form modelling program (GLASS), but
\phil{enables the input to this code to be provided in a novel way,}
via a user-provided diagram that is essentially a sketch of an
arrival-time surface.

An implementation \phil{of this method, \spl} is tested in a modelling
challenge using 29 simulated lenses \phil{drawn from the} set created
for \phil{the \sw citizen-science lens search}.  We find that
\phil{volunteers from this online community} asserted the image
parities and time ordering consistently in some lenses, but made
errors in other lenses depending on the image morphology. \phil{While}
errors in image parity and time ordering lead to  large errors in the
mass distribution, the enclosed mass \phil{was found to be more
robust: the model-derived Einstein radii found by the volunteers were
consistent with those produced by one of the professional team,
suggesting that given the appropriate tools, gravitational lens
modelling is a data analysis activity that can be crowd-sourced.}
Shortcomings \phil{of the implementation itself include} (a)~lens
profiles that tend to be systematically shallower than the lenses
\phil{(leading to systematic over-estimation of the Einstein radius 
by some $\sim25\%$}, and (b)~lack of detailed \phil{arc} modelling. 
Ideas for improvement are discussed.

\end{abstract}
